How to Analyze Build Logs

### 2021-04-19 tse-paper/paper.tex
start re-ordering chunk retrieval techniques section

\definecolor{todo}{HTML}{00a3cc}
% \newcommand{\todo}[1] {{\bfseries \color{todo} #1}}
% \newcommand{\caro}[1] {\todo{[Caro: #1]} } 
% hide todos:
\newcommand{\todo}[1] {}
\newcommand{\caro}[1] {}

@@

\section{Introduction}\label{sec:introduction}
In the past two decades, Continuous Integration (CI) has become a
ubiquitous best practice to streamline the build process of software
projects~\citep{hilton2016usage,beller2017oops,staahl2014modeling}.
Build logs are a textual by-product that these automatic software
builds create.
As a treasure trove of information~\citep{meyer},
build logs contain trace outputs of not only the elementary steps to
``make'' a project---such as retrieving and resolving dependencies and
compiling---but also about ancillary quality assurance steps---such as
testing the software and running automated static analysis
tools~\citep{beller2017oops}.
The
contents and format of build logs can vary highly depending on which tools
are involved in the build process, how they are
configured, and as projects and build tools evolve over
time~\citep{staahl2014modeling}.
Typically, though, they are at least a semi-structured and
somewhat stable journal of the commands executed during the build and
their results.

@@

The literature mapping shows the need for
build log
analysis techniques that are \emph{automated},
and can be \emph{configured} to target an \emph{arbitrary build tool}
with \emph{little effort} from the user.

@@

In this section, we formally define the concept of chunk
retrievals from build logs.
Then, we motivate and present the choice of three such techniques that
we implemented and evaluated for this study.

\subsection{Characteristics of Chunk Retrieval Techniques}
\label{sec:crt-characteristics}

With the term ``chunk retrieval technique,'' we refer to methods that
automatically extract strings that appear literally in build
logs, i.e., techniques that do not aggregate, combine, or deduct
information.
We call such pieces of information in build logs
\emph{chunks}.
Depending on the format of the chunks within the build log we
separate them into different \emph{structural categories}.
\todo{is it necessary to introduce strcutural categories here?
if yes, we should define it and give an example}

Generally, chunk retrieval techniques
can also be used to classify build logs, e.g.\, by checking
which substring was retrieved.
Therefore, they
cover the great majority of the types of log analysis we saw in
\Cref{sec:lit-sur:discussion}.

The techniques we investigate here do not require a formal lexer and
parser to analyze the entire structure of build logs, but focus on
ad hoc extraction of just one specific information piece per
configuration.

\todo{incorporate that this is done to adapt the chunk retrieval
technique for a certain build tool or other kind of information}
With the term \textit{configuration}, we abstract over the training
and parametrization that different techniques require in different
forms.
A configuration can be explicitly stated or implicitly derived
by learning through provided examples.
It is therefore a manual
specification of which information the chunk retrieval should target.
It also supplies the necessary information for the technique to
identify the targeted chunk in a build log.
Each chunk
retrieval technique has a specific \textit{granularity}, i.e.,\ the
smallest piece of text it can return (e.g., a line, or a word).
% The
% granularity might be adjustable by configuration.

\emph{Running a
chunk retrieval technique} means
executing a fully configured
technique to a build log.
The technique then outputs a (possibly empty) list of substrings of the
build log text.

@@

@@

The goal of this paper is to support practitioners and researchers in
their decision on how to analyze build logs.
To this end, we
conducted a systematic mapping to investigate how researchers
currently analyze build logs.
We saw that build log analysis is an emerging field.
At the same time, articles largely do not justify why
they use a certain technique, nor do they document it in detail or
open-source its implementation.
Moreover, the mapping showed that,
while there is a niche for custom parsers and plugins for very popular
tools, such as the Java ecosystem, there is a wide variety of build
tool and environment combinations for which the creation of such
bespoke tooling is unlikely.
Overall, this study showed the need for a new
\todo{not generic, but configurable}
generation of generic and automated build log analysis techniques that
need little effort to set up, both for practitioners as well as
researchers.

>>>>

\definecolor{todo}{HTML}{00a3cc}
\newcommand{\todo}[1] {{\bfseries \color{todo} #1}}
\newcommand{\caro}[1] {\todo{[Caro: #1]} } 
% hide todos:
% \newcommand{\todo}[1] {}
% \newcommand{\caro}[1] {}

@@

\section{Introduction}
\label{sec:introduction}
In the past two decades, Continuous Integration (CI) has become a
ubiquitous best practice to streamline the build process of software
projects~\citep{hilton2016usage,beller2017oops,staahl2014modeling}.
Build logs are a textual by-product that these automatic software
builds create.
As a treasure trove of information~\citep{meyer},
build logs contain trace outputs of not only the elementary steps to
``make'' a project---such as retrieving and resolving dependencies and
compiling---but also about ancillary quality assurance steps---such as
testing the software and running automated static analysis
tools~\citep{beller2017oops}.
The
contents and format of build logs can vary highly depending on which tools
are involved in the build process, how they are
configured, and as projects and build tools evolve over
time~\citep{staahl2014modeling}.
Typically, though, they are at least a semi-structured and
somewhat stable journal of the commands executed during the build and
their results.

@@

% \caro{can we soften this to lead into the new arguments in the following section?}
% The literature mapping shows the need for
% build log
% analysis techniques that are \emph{automated},
% and can be \emph{configured} to target an \emph{arbitrary build tool}
% with \emph{little effort} from the user.

@@

Based on our observations from the literature mapping, we believe that
researchers and developers would benefit from pre-made
build log analysis tools they
can re-use and configure to fit the kind of build logs and information
they are targeting.
Such \emph{configurable build log analysis tools} would save high effort
of developing custom tools we observed, and, when openly available,
improve the reproducibility of studies that mine information from
CI build logs.

To make a first step, this paper investigates three different approaches
to realize configurable build log analysis.
One of our central requirements is, that a future user should be able
to adapt the analysis with little effort to their use case.

To enable configurable build log analysis, we evaluate the following
techniques:
\begin{itemize}
  \item \textbf{\emph{Programming by Example (PBE)}}, which learns
  regular expressions based on few user-given log excerpts.
  We rely on the PROSE Framework \caro{link!} that is also responsible for auto-filling
  cells in Microsoft Excel.

  \item Inspired by information retrieval, a
  \textbf{\emph{Common Text Similarity (CTS)}} approach that finds log
  parts with similar vocabulary than that of user-given log excerpts.

  \item A \textbf{\emph{Keyword Search (KWS)}} that searches for terms
  provided by the user.
\end{itemize}

All these three techniques retrieve \emph{chunks} of text from an
analyzed build log, which is why we will further call them
\emph{Chunk Retrieval Techniques}.

In this section, we introduce these three techniques with the help of an
example, discuss their realization in our prototypes in more detail
and discuss the option of using machine learning to create a configurable
build log analysis approach.

% \subsection{Characteristics of Chunk Retrieval Techniques}
% \label{sec:crt-characteristics}

% With the term ``chunk retrieval technique,'' we refer to methods that
% automatically extract strings that appear literally in build
% logs, i.e., techniques that do not aggregate, combine, or deduct
% information.
% We call such pieces of information in build logs
% \emph{chunks}.
% Depending on the format of the chunks within the build log we
% separate them into different \emph{structural categories}.
% \todo{is it necessary to introduce strcutural categories here?
% if yes, we should define it and give an example}

% Generally, chunk retrieval techniques
% can also be used to classify build logs, e.g.\, by checking
% which substring was retrieved.
% Therefore, they
% cover the great majority of the types of log analysis we saw in
% \Cref{sec:lit-sur:discussion}.

% The techniques we investigate here do not require a formal lexer and
% parser to analyze the entire structure of build logs, but focus on
% ad hoc extraction of just one specific information piece per
% configuration.

% \todo{incorporate that this is done to adapt the chunk retrieval
% technique for a certain build tool or other kind of information}
% With the term \textit{configuration}, we abstract over the training
% and parametrization that different techniques require in different
% forms.
% A configuration can be explicitly stated or implicitly derived
% by learning through provided examples.
% It is therefore a manual
% specification of which information the chunk retrieval should target.
% It also supplies the necessary information for the technique to
% identify the targeted chunk in a build log.
% Each chunk
% retrieval technique has a specific \textit{granularity}, i.e.,\ the
% smallest piece of text it can return (e.g., a line, or a word).
% % The
% % granularity might be adjustable by configuration.

% \emph{Running a
% chunk retrieval technique} means
% executing a fully configured
% technique to a build log.
% The technique then outputs a (possibly empty) list of substrings of the
% build log text.

@@

\caro{To investigate which of the three techniques are feasible to
facilitate configurable build log analysis, we conduct an empirical
comparison on a large dataset of real world log files and excerpts.
}

@@

The goal of this paper is to support practitioners and researchers in
their decision on how to analyze build logs.
To this end, we
conducted a systematic mapping to investigate how researchers
currently analyze build logs.
We saw that build log analysis is an emerging field.
At the same time, articles largely do not justify why
they use a certain technique, nor do they document it in detail or
open-source its implementation.
Moreover, the mapping showed that,
while there is a niche for custom parsers and plugins for very popular
tools, such as the Java ecosystem, there is a wide variety of build
tool and environment combinations for which the creation of such
bespoke tooling is unlikely.
Overall, this study showed the need for a new
\todo{not generic, but configurable}generation of generic and automated build log analysis techniques that
need little effort to set up, both for practitioners as well as
researchers.
